---
date: "2019-02-17"
title: Notable Trees in Auckland, NZ
---
Here's an interactive map of 'notable' trees in the Auckland region I created this using [Leaflet](https://rstudio.github.io/leaflet/), a powerful mapping package in R. Take your time to scroll around the Auckland region and click on the green points to see what kind of special trees you have in your neighbourhood! ([datasource](https://koordinates.com/layer/6392-pup-notable-trees/))

<iframe width="100%" height="450" name="iframe" src="/trees.html" frameborder="0" scrolling="no" onload="resizeIframe(this)"></iframe>

The inspiration for this post came from [Isaac Bain's tutorial](http://www.isaacbain.co.nz/tutorials/GIS-R.html) where he used freely available data to map the DOC huts around NZ. You can re-create a similar visualisation with basically any shapefile by following the following steps:
```{r, eval=FALSE}
#load required libraries, set working directory and unzip the shapefile
library(tidyverse)
library(sp)
library(leaflet)
library(spdplyr)
library(rgdal)
setwd("C:\\Users\\user\\Documents\\trees")
unzip("kx-pup-notable-trees-SHP.zip")
```
I downloaded my data from [Koordinates](https://koordinates.com/) which holds GIS datasets from all around the world! In the next step I am reading in the data and converting it to the wgs84 standard.
```{r, eval=FALSE}
#read in dataset and transform to wgs84 standard
trees <- readOGR("C:\\Users\\user\\Documents\\trees","pup-notable-trees")
wgs84 = '+proj=longlat +datum=WGS84'
trees <- spTransform(trees, CRS(wgs84))
```
In this step I am creating the HTML popup when a user clicks on the data point. You can get pretty fancy with these but the shapefile I had did not have much other information about the tree apart from the name.
```{r, eval=FALSE}
#create pop-up info
tree_popup <- paste0("<strong>Name: </strong>", trees$NAME)
```
And here is where we see the beauty of Leaflet, in just a few lines of code we have created a nice map for users to explore.
```{r, eval=FALSE}
#Map notable trees, add circle marker, centre the map and change the map type
leaflet(trees) %>%
  
   addCircleMarkers(stroke = FALSE, radius = 3.5, color = "Green", fillOpacity = 0.7, popup = tree_popup, label=trees$NAME, labelOptions = labelOptions(textOnly = TRUE)) %>%
  
   setView(lng = 174.8335603 , lat = -36.8818881, zoom = 12) %>%  # centered around Auckland
  
   addTiles() %>%
  
   addProviderTiles("Stamen.Toner", group = "Toner") %>% #changing the map=type
  
   addProviderTiles(providers$Stamen.Terrain, group = "Terrain")
```



